Comparing vector-based and ACT-R memory models using large-scale datasets:\protect\\User-customized hashtag and tag prediction on Twitter and StackOverflow

### 2015-05-27 writeups/psycMethods/psycMethods.tex
foo

\abstract{
  The growth of social media and user-created content on online sites provides unique opportunities to study models of declarative memory.
  The tasks of choosing a hashtag for a tweet and tagging a post on StackOverflow were framed as declarative memory retrieval problems.
  Two state-of-the-art cognitively-plausible declarative memory models were evaluated on how accurately they predict a user's chosen tags: an ACT-R based Bayesian model and a random permutation vector-based model.
  Millions of posts and tweets were collected, and both declarative memory models were used to predict Twitter hashtags and StackOverflow tags.
  The results show that past user behavior of tag use is a strong predictor of future behavior.
  Furthermore, past behavior was successfully incorporated into the random permutation model that previously used only context.
  Also, ACT-R's attentional weight term was linked to a common entropy-weighting natural language processing method used to attenuate low-predictor words.
  Word order was not found to be strong predictor of tag use, and the random permutation model performed comparably to the Bayesian model without including word order.
  This shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
  Finally, model accuracy was moderate to high for the tasks, which supports the theory that choosing tags on StackOverflow and Twitter is primarily a declarative memory retrieval process. 
  The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
  and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
}

@@

The ACT-R cognitive architecture \parencite{Anderson2007} contains one of the leading theories of human cognitive processing.
The declarative memory component in ACT-R is a Bayesian statistical model.
This is the same type of co-occurrence-based Naive Bayes statistical model that is prominent in the more general machine learning field.
When these Bayesian models are used for machine learning, it is common to use a large number of observations to construct the co-occurrence matrix that is the foundation for the model.
The primary reason for this is that the stability of statistical models increase as sample size increases.

However, it is much less common in the cognitive modeling community to use a large number of observations to build up declarative memory when creating ACT-R models.
If a cognitive modeler is using the declarative memory component of ACT-R,
it is common to either construct the co-occurrences by hand, or use a small training set of data to initially build up the declarative memory system.
This approach can and has worked well.
Essentially, the modeler is only including the observations in declarative memory that are relevant to the specific task that they are studying.

@@

A few researchers have started to use these large databases to evaluate how well declarative memory models scale,
for both ACT-R Bayesian models \parencites{Fu2007,Pirolli2003,Stanley2013,Douglass2010} and vector-based memory models \parencites{Jones2007,Rutledge2008,Recchia2010,Sahlgren2008}.

@@

The core question motivating this research is:
What kinds of cognitively-plausible models can best predict the tags that a user on a social media site uses?
This question can be tested by generating a prediction for the most likely used tag whenever a user is about to generate a tag, and then comparing the user's chosen tag to the model's prediction.
This process can be framed as a memory retrieval problem.
The process of suggesting a relevant tag to a user can be thought of as a memory retrieval request for a tag, given prior tag use and current context.

Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010} and StackOverflow tag use \parencites{Stanley2013}.
Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
At least two of the current memory retrieval models are based on a co-occurrence methodology:
ACT-R's declarative memory (DM) retrieval theory and random permutation vector-based memory systems.

@@

These two memory systems were compared across two domains where users generate tags for posts: Twitter and StackOverflow.
Twitter is a microblogging service where users create 140 character ``tweets'' and broadcast those messages out to the users who follow their posts.
A user is free to annotate important concepts or words in the tweet by preceding words with ``\#,'' thereby making the word into a hashtag.
After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
Some recent and often used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.

The two memory retrieval models were also tested on tag use for StackOverflow posts.
StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
A user with a programming-related question creates a post with their question and tags the post with a few specific programming-related tags. 
The tags chosen by a user on the StackOverflow site represent the primary topic of the question, such as a specific programming language, tool, software package, or framework.
Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, and \emph{Common Lisp}.

@@

The ACT-R declarative memory theory and the vector-based retrieval models will be described.
The Latent Semantic Analysis (LSA) theory is closely related to vector-based retrieval models, so an overview of LSA will also be included.
The theories underlying the ACT-R and vector-based models will be compared.
Also, different tagging models have already been applied to various recommendation systems to tag user-generated content.
Many of these models are based on the same Bayesian statistical framework that is the foundation of the ACT-R declarative memory theory.
So these systems will be described and the models used will be discussed within the context of Bayesian statistical models, LSA approaches, and the ACT-R declarative memory theory.

@@

ACT-R \parencite{Anderson2007} is a cognitive architecture that formalizes how each cognitive process of the mind (e.g., memory, learning, visual and motor) interacts to produce behavior.
The declarative memory system is a component of that architecture that models the timing, learning, and forgetting processes that make up declarative memory storage and retrieval.
The equations that make up this system were derived from a rational analysis of declarative memory retrieval.
That is, given the task of retrieving a chunk of information from declarative memory,
the current context (i.e., external and internal environment state), and past experience (i.e., prior memories and exposure), 
what is the optimal behavior (i.e., the optimal chunk to retrieve from memory)?
Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of needing retrieval again, given the prior history of exposure to the chunk.
These chunk prior probabilities are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and adjusted by current environment state.

@@

Latent Semantic Analysis \parencite{Landauer1997} (LSA) is an alternative representation of declarative memory.
Both ACT-R's Bayesian declarative memory model and LSA compute activations based on a co-occurrence matrix.
However, LSA performs further processing on that matrix in order to collapse words that have similar meaning.
LSA uses the Singular Value Decomposition process to reduce the dimensionality of the word by document co-occurrence matrix, while maintaining as much of the variance in the original full matrix as possible.
The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.

@@

Models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites.
Google has even deployed a hashtag recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
A tag-suggestion tool has also been deployed for the \url{tex.stackexchange.org} and \url{superuser.com} StackExchange sites.
Models for Twitter and StackOverflow have been proposed and tested, but none have yet been deployed.
Recommendation models for StackOverflow and Twitter will be described in more detail to follow.

@@

The offset term was added to normalize the mean activation across retrievals.
This allowed the use of a logistic regression statistical technique to calibrate the weights for each of the terms.
The scaled entropy measure $E_{j}$ was chosen for the attentional weight ($W_{j}$) term because it has a strong theoretical foundation \parencite{Dumais1991} and is parameter free.

@@

Several hashtag recommendation models for Twitter have been developed recently.
One of the first models used a Bayesian co-occurrence statistical technique to predict the most likely hashtag associated with a tweet as a function of prior hashtag use and context \parencite{Mazzia2009}.
The model presented by \textcite{Mazzia2009} is quite similar to an ACT-R declarative retrieval model.
The main potential difference is that --- like \textcite{Stanley2013} --- the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.

Several other models have taken a tweet-centered approach to hashtag prediction, where suggested hashtags are collected from hashtags used from similar tweets.
\textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurrence-based similarity score between a composed tweet and the rest of the tweets in the database.

There has been a mix of co-occurrence models created to predict hashtag use on Twitter, and the results are encouraging.
However, there has been little research on how a specific user's past hashtag use should influence the model's prediction when that user is composing a tweet.
\textcite{Kywe2012} did take a user-centered view with their model, but the hashtags predicted by the user and content were analyzed separately and then the top from each group were combined for prediction.
The ACT-R declarative memory retrieval theory shows how past user behavior and current context combine to produce the most likely retrieval.
Each component is summed together to generate a total activation, and then the likelihood of chunks are ranked by that summed activation.
This technique is compensatory and allows for a more natural combination of the two components, where weights can be assigned to each component depending on how strongly each term predicts performance.

@@

Unlike StackOverflow, the entire Twitter dataset is not released to the public.
Due to this and the scale of Twitter, only a part of the Twitter dataset relevant to this research was collected.

@@

Four popular-hashtags datasets were collected at four different time periods, spanning across one month.
Multiple datasets were collected to ensure that the results were not dependent on a particular sample of hashtags and would generalize across different popular-hashtag datasets and different time periods.
Each dataset required around four days to collect enough tweets to reach the quota, and one collection was done each week.

@@

The entire StackOverflow dataset is packaged and released quarterly to the public.
So the most recent packaged dataset was downloaded and the database tables were imported into PostgreSQL.

@@

Twitter provides two types of APIs for acquiring data: one based on streaming and a direct query-based interface.
The Tweepy Python package provides a Python functional interface to these APIs, and this package was used to collect data from Twitter.
The query-based interface was used to collect all tweets for specific popular users.
At most \num{3000} of the most recent tweets were collected for each user, since Twitter only allows access to this number of tweets for each user.

@@

A custom tokenization algorithm created by \textcite{Owoputi2013} was used to chunk the words in the text.
This algorithm was specifically created for tokenizing Twitter content.
However, during testing it was also found to be robust and well suited for StackOverflow content.
The initial model results with the new tokenizer were qualitatively similar to the model results found in \textcite{Stanley2013} where the Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} was used.
So the same tokenization algorithm from \textcite{Owoputi2013} was used for both StackOverflow and Twitter, since performance was similar to \textcite{Stanley2013} for StackOverflow,
and it seemed beneficial to use the same algorithm for both datasets to make comparisons across datasets more equitable.

@@

The StackOverflow and Twitter popular-users datasets were used to characterize tag growth and decay for a user on each site.
The popular-users dataset contains 12 slices of popular users for both StackOverflow and Twitter.
Each user's tag behavior from the popular-users dataset for each site was analyzed.

22 different values of the decay rate parameter were explored for each slice
(0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6, 1.8, 2, 5, 10, 20).
The values were chosen to look at how performance changes from a pure frequency model ($d=0$), to a blend of frequency and recency ($d \approx .5$), to a pure recency model ($d=20$).
The decay rate value was varied for both the standard form and optimized learning form of $B_{i}$. 
Only values equal to or below 1 were examined for the optimized learning form, since values above 1 for this form cause a bifurcation and are not psychologically interesting.
The bifurcation occurs because at a decay rate of 1 the denominator of the first term to compute $B_{i}$ goes to zero (see Table \ref{tabACTRBLLModel}).

@@

Model accuracies at each decay rate value for each user in the dataset slice are included in the plots.
Each user's mean accuracy was subtracted off in order to compute a normalized mean for that user.
This was done so that the relative accuracy change for different decay rate values could be more easily visualized.
The plotting technique is a analogous to what is done when testing the main effect in a repeated measures design (i.e., each subject's mean score is subtracted off).

@@

This section will examine how model accuracy is influenced by four architectural concerns:
method of handling stop words, influence of dataset size, method to combine prior and context, and word order.
The four architectural concerns were examined for the following reasons:
During exploratory evaluation, model accuracy seemed to be highly sensitive to the method of dealing with the commonly-used stop words, so several different methods were examined in more detail.
Due to research from \textcite{Budiu2007} and \textcite{Recchia2010}, model accuracy for the Bayesian and vector-based models is influenced by the number of documents in the dataset,
so the effect of dataset size on accuracy was examined in more detail.
One of the primary goals of this research is to identify the best method to combine the model term for past user behavior with the context term for the random permutation model,
so two different ways to add the model terms were tested.
Finally, previous research has shown that including word order in the vector-based models improves performance \parencites{Jones2007, Sahlgren2008},
so word order was also examined to see if the effect could be replicated and also to measure the overall importance of word order on model accuracy relative to other architectural manipulations.

@@

\textcite{Stanley2013} used the entropy weighting technique recommended in \textcite{Dumais1991} for the StackOverflow dataset.
The result maps cleanly to the ACT-R DM theory by modifying each word's attentional weight ($W_{j}$) parameter.
It also seems less arbitrary than explicitly identifying a list of stop words to remove from the analysis.

@@

The runs from the Twitter popular-hashtags and StackOverflow randomly-sampled datasets were used for the analysis.
Both the random permutation and Bayesian models were tested.
Activations for context were computed by using five different methods to handle stop words:
The two removal methods used in \textcite{Sahlgren2008}
(i.e., removal based on frequency, removal based on the pre-determined 571-word Cornell SMART stoplist \parencite{Salton1988}),
the entropy-weighting method used in \textcite{Stanley2013},
combining the entropy-weighting and frequency-filtering methods,
and no removal or weighting of stop words.

@@

The no processing plot shows the size of the spikes when high-frequency words are not removed.
The entropy plot shows how the size decreases by more than an order of magnitude after the entropy weighting measure is used.
The cutoff for the frequency plot was chosen such that the plot looks similar to the entropy plot, with spikes around the same size.
The cutoff that produced the frequency plots in Figures \ref{figContextCutoffSO} and \ref{figContextCutoffT} is where a word represents more than \num{.04}\% of total occurrences in the dataset.
This may or may not be equal to the value for \textcite{Sahlgren2008}, since that number was expressed in total observations (not as a percentage), and the total number of words in the dataset was not given.

@@

\textcite{Stanley2013} found that the entropy-weighting technique to attenuate stop words for the Bayesian model significantly improved model accuracy.
The same effect is found for the StackOverflow and Twitter datasets in this study.
Further, when this entropy-weighting technique is compared directly to the more common frequency-filtering technique, weighting by entropy produces more accurate results.
This was clearly the case for the Bayesian model.
Both frequency and entropy techniques performed equally for the random permutation model.
However, since the entropy-weighting technique is parameter free and does not require a cutoff to tune, it is also a better choice for the random permutation model for these datasets.

@@

It is interesting to compare model accuracy between the Bayesian and random permutation model at the smallest and largest corpus sizes for both StackOverflow and Twitter.
At smaller corpus sizes, the random permutation model outperforms the Bayesian model (slightly for Twitter, dramatically for StackOverflow), and this performance difference deteriorates as corpus size increases.
In other words, the Bayesian model needs larger co-occurrence matrices than the random permutation model to work properly.
As the size of the dataset increases, the compression may start to lose crucial information, and the uncompressed Bayesian model should be used.
As the size of the dataset decreases, the noise and crosstalk introduced by the compression for the random permutation model is actually beneficial, and the compressed random permutation model should be used. 
This noise is beneficial because it helps soften spurious $S_{ji}$ values when the counts for each individual cell in the co-occurrence matrix are still volatile, which happens at smaller corpus sizes.

@@

The transition point where accuracy for the Bayesian model outperformed the random permutation model was much higher for StackOverflow than Twitter.
This is most likely because StackOverflow has almost \num{100} times more unique tags than the Twitter popular-hashtags dataset.
The number of words in a StackOverflow post is also much higher than Twitter, which should have lowered the transition point.
However, the effect of words in a post must be less than the effect of number of unique tags, as the transition point was still much higher for StackOverflow given that it should have lowered due to post size.

@@

As a final architectural concern, vector-based models with and without word order were tested to examine the effect of word order on model accuracy.
Vector-based models like the random permutation model can easily incorporate word order when making predictions.
This is one of their primary strengths that has been discussed \parencites{Jones2007, Sahlgren2008}.
In order to thoroughly test the impact of adding word order into the model, various methods for incorporating word order were tested on the StackOverflow and Twitter datasets.

@@

The general discussion is separated into two parts.
The first section will describe the overall results, reemphasize key findings, and discuss the theoretical implications of each key finding.
The final section will describe future applications and directions for the research.

@@

Overall, both the random permutation and Bayesian declarative memory models fit the observed tagging behavior in StackOverflow and Twitter relatively well.
The fits for StackOverflow in particular are quite high, especially given the large tag space on the site.
These results provide support for the idea that choosing a hashtag when composing a tweet and tagging a StackOverflow post is akin to a declarative memory retrieval process.
However, both models were not 100 percent accurate, and the amount of variance left to predict was much larger than what would be observed if retrieval noise was included in the model.
So it is most likely the case that the task is not done solely with declarative memory processes.

@@

These results pose an interesting and important fundamental question about declarative memory:
How, precisely, is the neural system configured to represent co-occurrences for declarative memory?
Since at the behavioral level both models perform similarly, it is possible that the neural system is using either representation to store co-occurrences.
Vector-based models are simple enough that they can easily be implemented in a neural network.
Perhaps a form of vector-based compressed storage is being used at the neural level,
since it can be implemented in neurons and is an efficient way for the machinery to generate predictions that closely match the statistical structure of the environment
(i.e., that closely match the uncompressed Bayesian representation).

@@

The relatively strong performance of the random permutation model is not due to the model's ability to represent word order.
Performance is similar for the random permutation and Bayesian models when the co-occurrence matrix for both models is the same, and not based on word order.
Performance also does not improve by much when word order is added as the last predictor in the random permutation model.
The random permutation model seems to work so well primarily due to the way that the full co-occurrence representation is compressed.

@@

\subsubsection{Future Applications}

>>>>

\abstract{
  The growth of social media and user-created content on online sites provides unique opportunities to study models of human declarative memory.
  The tasks of choosing a hashtag for a tweet and tagging a post on StackOverflow were framed as declarative memory retrieval problems.
  Two state-of-the-art cognitively-plausible declarative memory models were evaluated on how accurately they predict a user's chosen tags: an ACT-R based Bayesian model and a random permutation vector-based model.
  Millions of posts and tweets were collected, and both declarative memory models were used to predict Twitter hashtags and StackOverflow tags.
  The results show that past user behavior of tag use is a strong predictor of future behavior.
  Furthermore, past behavior was successfully incorporated into the random permutation model that previously used only context.
  Also, ACT-R's attentional weight term was linked to an entropy-weighting natural language processing method used to attenuate high-frequency words (e.g., articles and prepositions).
  Word order was not found to be strong predictor of tag use, and the random permutation model performed comparably to the Bayesian model without including word order.
  This shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
  Finally, model accuracy was moderate to high for the tasks, which supports the hypothesis that choosing tags on StackOverflow and Twitter is essentially a declarative memory retrieval process. 
  The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
  and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
}

@@

The ACT-R cognitive architecture \parencite{Anderson2007} contains one of the leading theories of human cognitive processing.
The declarative memory component in ACT-R is a Bayesian statistical model.
This is the same type of co-occurrence-based Naive Bayes statistical model that is prominent in the field of machine learning.
When these Bayesian models are used for machine learning, it is common to use a large number of observations to construct the co-occurrence matrix that is the foundation for the model.
The primary reason for this is that the stability of statistical models increase as sample size increases.

However, it is much less common in the cognitive modeling community to use a large number of observations to build up declarative memory when creating ACT-R models.
If a cognitive modeler is using the declarative memory component of ACT-R,
it is common to either construct the co-occurrences by hand, or use a small training set of data to initially parameterize the declarative memory system.
This approach can and has worked well.
Essentially, the modeler is only including the observations in declarative memory that are relevant to the specific task that they are studying.

@@

A few researchers have started to use large-scale databases to evaluate how well declarative memory models scale,
for both ACT-R Bayesian models \parencites{Fu2007,Pirolli2003,Stanley2013,Douglass2010} and vector-based memory models \parencites{Jones2007,Rutledge2008,Recchia2010,Sahlgren2008}.

@@

The core question motivating this research is:
What kinds of cognitively-plausible models can best predict the tags that a user on a social media site selects?
This question can be tested by generating a prediction for the most likely selected tag whenever a user is about to generate a tag, and then comparing the user's chosen tag to the model's prediction.
The process of suggesting a relevant tag to a user can be thought of as a memory retrieval request for a tag, given prior tag use and current context.

Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag selection \parencite{Efron2010} and StackOverflow tag selection \parencites{Stanley2013}.
Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
At least two of the current memory retrieval models are based on a co-occurrence methodology:
ACT-R's declarative memory (DM) retrieval theory and random permutation vector-based memory systems.

@@

These two memory models were compared across two domains where users generate tags for posts: Twitter and StackOverflow.
Twitter is a microblogging service where users create 140 character ``tweets'' and broadcast those messages out to the users who follow their posts.
A user is free to annotate important concepts or words in the tweet by preceding words with ``\#,'' thereby making the word into a hashtag.
After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
Some recent and often used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.

The two memory retrieval models were also tested on tag selection for StackOverflow posts.
StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
A user with a programming-related question creates a post with their question and tags the post with a few specific programming-related tags. 
The tags chosen by a user on the StackOverflow site represent the primary topic of the question, such as a specific programming language, tool, software package, or framework.
Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, and \emph{Common Lisp}.

@@

The ACT-R declarative memory model and vector-based retrieval model will be described.
Latent Semantic Analysis (LSA) is closely related to vector-based retrieval models, so an overview of LSA will also be included.
The theories underlying the ACT-R and vector-based models will be compared.
Also, different tagging models have already been applied to various recommendation systems to tag user-generated content.
Many of these models are based on the same Bayesian statistical framework that is the foundation of the ACT-R declarative memory theory.
So these systems will be described and the models used will be discussed within the context of Bayesian statistical models, LSA approaches, and the ACT-R declarative memory theory.

@@

ACT-R \parencite{Anderson2007} is a cognitive architecture that formalizes how each cognitive process of the mind (e.g., memory, learning, visual and motor) interacts to produce behavior.
The declarative memory system is a component of the architecture that models the timing, learning, and forgetting processes that make up declarative memory storage and retrieval.
The equations that make up this system were derived from a rational analysis of declarative memory retrieval.
That is, given the task of retrieving a chunk of information from declarative memory,
the current context (i.e., external and internal environment state), and past experience (i.e., prior memories and exposure), 
what is the optimal behavior (i.e., the optimal chunk to retrieve from memory)?
Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of needing retrieval again, given the prior history of exposure to the chunk.
These chunk prior probabilities are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and adjusted by current environment state.

@@

Latent Semantic Analysis \parencite{Landauer1997} (LSA) is an alternative representation of declarative memory.
Both ACT-R's Bayesian declarative memory model and LSA compute activations based on a co-occurrence matrix.
However, LSA performs further processing on that matrix in order to collapse words that have similar meaning.
LSA uses the Singular Value Decomposition (SVD) process to reduce the dimensionality of the word by document co-occurrence matrix, while maintaining as much of the variance in the original full matrix as possible.
The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.

@@

Models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites,
though most of these models are not based on theories of human declarative memory.
Google has even deployed a hashtag recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
A tag-suggestion tool has also been deployed for the \url{tex.stackexchange.org} and \url{superuser.com} StackExchange sites.
Models for Twitter and StackOverflow have been proposed and tested, but none have yet been deployed.
Recommendation models for StackOverflow and Twitter will be described in more detail to follow.

@@

The offset term was added to normalize the mean activation across retrievals.
This allowed the use of a logistic regression statistical technique to calibrate the weights for each of the terms.
The scaled entropy measure $E_{j}$ was chosen for the attentional weight ($W_{j}$) term because it has a strong theoretical foundation \parencite{Dumais1991} and is parameter-free.

@@

Several hashtag recommendation models for Twitter have been developed recently.
One of the first models used a Bayesian co-occurrence statistical technique to predict the most likely hashtag associated with a tweet as a function of prior hashtag use and context \parencite{Mazzia2009}.
This is quite similar to an ACT-R declarative retrieval model.
The main potential difference is that---like \textcite{Stanley2013}---the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.

Several other models have taken a tweet-centered approach to hashtag prediction, where suggested hashtags are collected from hashtags used in similar tweets.
\textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurrence-based similarity score between a composed tweet and the rest of the tweets in the database.

There has been a mix of co-occurrence models created to predict hashtag use on Twitter, and the results are encouraging.
However, there has been little research on how a specific user's past hashtag use should influence the model's prediction when that user is composing a tweet.
\textcite{Kywe2012} did take a user-centered view with their model, but the hashtags predicted by the user and content were analyzed separately and then the top from each group were combined for prediction.
The ACT-R declarative memory retrieval theory shows how past experience and current context combine to produce the most likely retrieval.
Each component is summed together to generate a total activation, and then the likelihood of chunks are ranked by that summed activation.
This technique is compensatory and allows for a more natural combination of the two components, where weights can be assigned to each component depending on how strongly each term predicts performance.

@@

Unlike StackOverflow, the entire Twitter dataset is not released to the public.
Due to this and the scale of Twitter, only subsets of the Twitter dataset were examined.
The relevant subsets are described below.

@@

Four popular-hashtags datasets were collected at four different time periods, spanning one month.
Multiple datasets were collected to ensure that the results were not dependent on a particular sample of hashtags and would generalize across different popular-hashtag datasets and different time periods.
Each dataset required around four days to collect enough tweets to reach the quota, and one collection was done each week.

@@

The entire StackOverflow dataset is packaged and released quarterly to the public.
So the most recent packaged dataset was downloaded and the tables were imported into a relational database.

@@

Twitter provides two types of APIs for acquiring data: one based on streaming and a direct query-based interface.
The query-based interface was used to collect all tweets for specific popular users.
At most \num{3000} of the most recent tweets were collected for each user, since Twitter only allows access to this number of tweets for each user.

@@

A custom tokenization algorithm created by \textcite{Owoputi2013} was used to chunk the words in the text.
This algorithm was specifically created for tokenizing Twitter content.
However, during testing it was also found to be robust and well-suited for StackOverflow content.
The initial model results with the new tokenizer were qualitatively similar to the model results found in \textcite{Stanley2013} where the Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} was used.
So the same tokenization algorithm from \textcite{Owoputi2013} was used for both StackOverflow and Twitter, since it seemed beneficial to use the same algorithm for both datasets to make comparisons across datasets more equitable.

@@

The StackOverflow and Twitter popular-users datasets were used to characterize tag growth and decay for a user on each site.
Each user's tag behavior from each site was analyzed.

22 different values of the decay rate parameter were explored
(0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6, 1.8, 2, 5, 10, 20).
The values were chosen to look at how performance changes from a pure frequency model ($d=0$), to a blend of frequency and recency ($d \approx .5$), to a pure recency model ($d=20$).
The decay rate value was varied for both the standard form and optimized learning form of $B_{i}$. 
Only values equal to or below 1 were examined for the optimized learning form, since values above 1 for this form cause a bifurcation and are not psychologically interesting.
The bifurcation occurs because at a decay rate of 1 the denominator of the first term to compute $B_{i}$ goes to zero (see Table \ref{tabACTRBLLModel}).

@@

Model accuracies at each decay rate value for each user in the dataset slice are included in the plots.
Each user's mean accuracy was subtracted in order to compute a normalized mean for that user.
This was done so that the relative accuracy change for different decay rate values could be more easily visualized.
The plotting technique is a analogous to what is done when testing the main effect in a repeated measures design (i.e., each subject's mean score is subtracted).

@@

This section will examine how model accuracy is influenced by four architectural concerns:
method of handling stop words, influence of dataset size, method to combine prior and context, and word order.
These particular architectural concerns were examined for the following reasons:
During exploratory evaluation, model accuracy seemed to be highly sensitive to the method of dealing with the commonly-used stop words, so several different methods were examined in more detail.
Due to research from \textcite{Budiu2007} and \textcite{Recchia2010}, model accuracy for the Bayesian and vector-based models is influenced by the number of documents in the dataset,
so the effect of dataset size on accuracy was examined in more detail.
One of the primary goals of this research is to identify the best method to combine the model term for past user behavior with the context term for the random permutation model,
so two different ways to add the model terms were tested.
Finally, previous research has shown that including word order in the vector-based models improves performance \parencites{Jones2007, Sahlgren2008},
so word order was also examined to see if the effect could be replicated and also to measure the overall importance of word order on model accuracy relative to other architectural manipulations.

@@

\textcite{Stanley2013} used the entropy weighting technique recommended in \textcite{Dumais1991} for the StackOverflow dataset.
This method maps cleanly to the ACT-R DM theory by modifying each word's attentional weight ($W_{j}$) parameter.
It also seems less arbitrary than explicitly identifying a list of stop words to remove from the analysis.

@@

The runs from the Twitter popular-hashtags and StackOverflow randomly-sampled datasets were used for the analysis.
Both the random permutation and Bayesian models were tested.
Activations for context were computed by using five different methods to handle stop words:
The two removal methods used in \textcite{Sahlgren2008}
(i.e., removal based on frequency, removal based on the pre-determined 571-word Cornell SMART stoplist: \textcite{Salton1988}),
the entropy-weighting method used in \textcite{Stanley2013},
combining the entropy-weighting and frequency-filtering methods,
and no removal or weighting of stop words.

@@

The ``no processing'' plot shows the size of the spikes when high-frequency words are not removed.
The entropy plot shows how the size decreases by more than an order of magnitude after the entropy weighting measure is used.
The cutoff for the frequency plot was chosen such that the plot looks similar to the entropy plot, with spikes around the same size.
The cutoff that produced the frequency plots in Figures \ref{figContextCutoffSO} and \ref{figContextCutoffT} is where a word represents more than \num{.04}\% of total occurrences in the dataset.
This may or may not be equal to the value for \textcite{Sahlgren2008}, since that number was expressed in total observations (not as a percentage), and the total number of words in the dataset was not given.

@@

\textcite{Stanley2013} found that the entropy-weighting technique to attenuate stop words for the Bayesian model significantly improved model accuracy.
The same effect is found for the StackOverflow and Twitter datasets in this study.
Further, when this entropy-weighting technique is compared directly to the more common frequency-filtering technique, weighting by entropy produces more accurate results.
This was clearly the case for the Bayesian model.
Both frequency and entropy techniques performed equally for the random permutation model.
However, since the entropy-weighting technique is parameter-free and does not require a cutoff to tune, it is also a better choice for the random permutation model for these datasets.

@@

It is interesting to compare model accuracy between the Bayesian and random permutation model at the smallest and largest corpus sizes for both StackOverflow and Twitter.
At smaller corpus sizes, the random permutation model outperforms the Bayesian model (slightly for Twitter, dramatically for StackOverflow), and this performance difference diminishes as corpus size increases.
In other words, the Bayesian model needs larger co-occurrence matrices than the random permutation model to work properly.
As the size of the dataset increases, the compression may start to lose crucial information, and the uncompressed Bayesian model should be used.
As the size of the dataset decreases, the noise and crosstalk introduced by the compression for the random permutation model is actually beneficial, and the compressed random permutation model should be used. 
This noise is beneficial because it helps soften spurious $S_{ji}$ values when the counts for each individual cell in the co-occurrence matrix are still volatile, which happens at smaller corpus sizes.

@@

The transition point where accuracy for the Bayesian model outperformed the random permutation model was much higher for StackOverflow than Twitter.
This is most likely because StackOverflow has almost \num{100} times more unique tags than the Twitter popular-hashtags dataset.
The number of words in a StackOverflow post is also much higher than Twitter, which should have lowered the transition point.
However, the effect of the number of words in a post must be less than the effect of number of unique tags, as the transition point was still much higher for StackOverflow given that it should have lowered due to post size.

@@

As a final architectural concern, vector-based models with and without word order were tested to examine the effect of word order on model accuracy.
Vector-based models like the random permutation model can easily incorporate word order when making predictions.
This has been touted as one of their primary strengths \parencites{Jones2007, Sahlgren2008}.
In order to thoroughly test the impact of adding word order into the model, various methods for incorporating word order were tested on the StackOverflow and Twitter datasets.

@@

The general discussion is separated into two parts.
The first section will describe the overall results, reemphasize key findings, and discuss the theoretical implications of each key finding.
The final section will describe potential applications and directions for the research.

@@

Overall, both the random permutation and Bayesian declarative memory models fit the observed tagging behavior in StackOverflow and Twitter reasonably well.
The fits for StackOverflow in particular are quite high, especially given the large tag space on the site.
These results provide support for the idea that choosing a hashtag when composing a tweet and tagging a StackOverflow post is akin to a declarative memory retrieval process.
However, both models were not 100 percent accurate, and the amount of variance left to predict was much larger than what would be observed if retrieval noise was included in the model.
One possibility that seems likely is that the tag selection process includes other processes besides just declarative memory retrieval
(e.g., strategic processes, determining the number of tags to employ, attempts to add humor).

@@

These results pose an interesting and important fundamental question about declarative memory:
How, precisely, is the neural system configured to represent co-occurrences for declarative memory?
Since at the behavioral level both models perform similarly, it is possible that the neural system is using something similar to either representation to store co-occurrences.
Vector-based models are simple enough that they can easily be implemented in a neural network.
Perhaps a form of vector-based compressed storage is being used at the neural level,
since it can be implemented in neurons and is an efficient way for the machinery to generate predictions that closely match the statistical structure of the environment
(i.e., that closely match the uncompressed Bayesian representation).

@@

The relatively strong performance of the random permutation model is not due to the model's ability to represent word order.
Performance is similar for the random permutation and Bayesian models when the co-occurrence matrix for both models is the same, and not based on word order.
Performance also does not improve by much when word order is added as the last predictor in the random permutation model.
The random permutation model seems to work so well mainly due to the way that the full co-occurrence representation is compressed.

@@

\subsubsection{Potential Applications}
